---
title: "Coctail reductions"
author: "Jonathan Speh"
date: "1/22/2021"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
suppressPackageStartupMessages({
  library(tidyverse)
  library(tidymodels)
  library(tidytext)
  library(embed)
})
```

# Load and prepare data

## Download and inspect
```{r download, message=FALSE}
cocktails <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-26/boston_cocktails.csv")

knitr::kable(cocktails%>%group_by(ingredient)%>%tally(sort=TRUE)%>%top_n(5), caption = "Tally of ingredients in the raw data")
```

## Wrangling

+ Remove inconsistet naming
+ parse ingredients into numeric
+ convert into units that were not invented by someone who already had a few of those cocktails to many
+ filter out ingredients that are used less than 15 times

### Parse
```{r wrangle, warning=FALSE}
cocktails%>%mutate(ingredient=str_to_lower(ingredient),
                   ingredient=str_replace(ingredient,"-"," "),
                   ingredient=str_remove(ingredient, " liquor"),
                   ingredient=str_remove(ingredient, "(if desired)"),
                   ingredient=case_when(str_detect(ingredient, "bitters") ~ "bitters",
                                        str_detect(ingredient,"orange") ~ "orange juice",
                                        str_detect(ingredient,"lemon") ~ "lemon juice",
                                        str_detect(ingredient,"lime") ~ "lime juice",
                                        str_detect(ingredient,"grapefruit") ~ "grapefruit juice",
                                        TRUE ~ ingredient),
                   measure=case_when(str_detect(ingredient, "bitters") ~ str_replace(measure, "oz$", "dash"),
                                     TRUE ~ measure),
                   measure = str_replace(measure, " ?1/2",".5"),
                   measure = str_replace(measure, " ?3/4",".75"),
                   measure = str_replace(measure, " ?1/4",".25"),
                   measure_number = parse_number(measure),
                   measure_number = if_else(str_detect(measure, "dash$"),
                                            measure_number / 50,
                                            measure_number),
                   measure_number=measure_number * 29.57) %>%
  add_count(ingredient)%>%
  filter(n > 15) %>%
  select(-n)%>%
  distinct(row_id,ingredient, .keep_all = TRUE)%>%
  na.omit()->cocktails_parsed

knitr::kable(head(cocktails_parsed), caption = "Cocktail data after tidying, filtering and parsing")
```

### bring in long format
```{r reshape}
cocktails_parsed%>%select(-ingredient_number, -row_id, -measure)%>%
  pivot_wider(names_from = ingredient, values_from=measure_number, values_fill = 0)%>%
  janitor::clean_names()->cocktails_wide

knitr::kable(head(cocktails_wide), caption= "Cocktail data in wide format")
```

# Dimension reduction
## Principle component analysis

+  Unsupervicesd clustering using the tidymodels package. 
+  Ingredients are predictors
+  `all_predictors()` selects variables that have no role
+  `recipe()`function sets up (explains) analyses, does not compute anything
+  `prep` estimates required parameters from training set 


```{r prepPCA}
recipe(~ ., data=cocktails_wide)%>%
  update_role(name, category, new_role = "id")%>%
  step_normalize(all_predictors())%>% # centers and scales all predictors, i.e all ingredients
  step_pca(all_predictors())->pca_rec

prep(pca_rec)->pca_prep

knitr::kable(pca_prep[2], caption = "Assignment of the variables after PCA")
```
### tidy PCA and inspect it

```{r tidyPCA, fig.cap="Barlot showing the contributions of the different ingredients on the first five components"}
tidy(pca_prep, 2)->pca_tidy # selects prep

pca_tidy%>%
  filter(component %in% paste0("PC",1:5))%>%
  ggplot(aes(value,terms,fill=terms))+
  geom_col(show.legend = FALSE)+
  facet_wrap(~component, nrow = 1)
```
+ Plot shows how much the components contribute to the different components. 
  + Bars in different directions --> either or in the drink 

```{r, fig.cap="Simplified barplot of the first four components. The bars schow the absolute contributionm colours indicated wether positive or negative"}
pca_tidy%>%
  filter(component %in% paste0("PC",1:4))%>%  # Select components 1:4
  group_by(component)%>%
  top_n(8, abs(value))%>%  # select top8 contributors for each component (by absolute value)
  ungroup()%>%
  mutate(terms=reorder_within(terms,abs(value), component))%>%  #order by contribution
  ggplot(aes(abs(value),terms,fill=value>0))+    # value > 0 explains positve or negative contribution
  geom_col()+
  scale_y_reordered()+
  facet_wrap(~component, scales = "free_y")
```
    
+  PC1 explains most variante
  +  powdered sugar vs simple sirup makes largest difference
  +  sweed / egg gine vs. sirup, lime tequilla
+  PC2 
  +  Vermouth vs lemony

### juice PCA
+  by default - 5 compnents
    +  reduces 40 dimensions to 5 
```{r juice, fig.cap="Dotplot of the first two components"}
juice(pca_prep)%>%
  ggplot(aes(PC1,PC2, label=name))+
  geom_point(aes(colour = category), alpha=0.7,size=2)+
  geom_text(check_overlap = TRUE, hjust="inward")+
  labs(colour = NULL)

```
+ First two components explain most variance in the ingredients



## UMAP 
+ unsupervised ML algorithm for dimension reduction
+ tidymodels does not need much changes in the input 

```{r prepUMAP}
recipe(~ ., data=cocktails_wide)%>%
  update_role(name, category, new_role = "id")%>%
  step_normalize(all_predictors())%>% # centers and scales all predictors, i.e all ingredients
  step_umap(all_predictors())->umap_rec

prep(umap_rec)->umap_prep

knitr::kable(umap_prep[2],caption="Results of UMAP computation")
```
### Juice UMAP
```{r juiceUMAP, fig.cap="UMAP plot indicating similarity of the different cocktails."}
juice(umap_prep)%>%
  ggplot(aes(umap_1,umap_2, label=name))+
  geom_point(aes(colour = category), alpha=0.7,size=2)+
  geom_text(check_overlap = TRUE, hjust="inward")+
  labs(colour = NULL)
```




